---
title: 'Intro - Spring 2020'
author: 
- name: Sathwik Kesappragada
  email: 
output:
  html_document:
    toc: yes
---

***

**Load necessary packages**

```{r, collapse=T}
library(datasets) # for the `quakes` data set

## You need to install the maps package first, then you can comment out the following line.
# install.packages("maps")

library(maps) # for map visualization
```


***

The `quakes` data set give the locations of 1000 seismic events of MB > 4.0. The events occurred in a cube near Fiji since 1964.

\footnotesize
```{r}
library(datasets) # we have already loaded the package but we can do it again
?quakes 
```

```{r, collapse=T}
class(quakes)
head(quakes, n=5) # print first 5 rows of quakes
dim(quakes)  # dimension of the table
names(quakes)  # list the variables in quakes
str(quakes)  # list the structures in quakes
```



```{r, collapse = T}
par(mfrow=c(2,2))# keeps the subfigures in a 2 by 2 layout
hist(quakes$mag, freq=F, xlab="Earthquake Magnitude") # histogram - subfigure 1
boxplot(quakes$mag, add=F, horizontal=T, xlab="Magnitude",  at=10, width=5, col=2, border="blue") #subfigure 2
lines(density(quakes$mag), col="red", lwd=3) # empirical density curve - subfigure 1
plot.ecdf(quakes$mag, verticals=T, pch="",
          xlab="Earthquake Magnitude", 
          main="ECDF of Earthquake Magnitudes") # empirical CDF plot - subfigure 3
qqnorm(quakes$mag)
qqline(quakes$mag,lwd=4, lty=4, col="blue") #subfigure 4 normal qq plot

```


There are several outliers labeled by open circles in the boxplot. Write a simple code to print out the indexes of the outliers and then print out the outlier observations.  


```{r, collapse=T}
x <-quakes$mag #store all values of magnitude vector 
summary(x)
q1 <- summary(x)[2] #get the second element of the summary list
q3 <- summary(x)[5]
iqr <- q3-q1
lwr <- q1 - 1.5*iqr #lowerbound
lwr
upr <- q3 + 1.5*iqr #upperbound
upr
which(x < lwr) # spits out the index if there exists outliers below the lowerbound
which(x > upr) #upperbound outlier indexes
c(x[x < lwr], x[x > upr]) #prints the values of the outlier 
```
How many outliers have you found?
**7**


That is, $X_i \sim N(\mu, \sigma^2)$, where $X_i$ is the magnitude of the $i$-th earthquake observation and $i=1,2,\cdots,1000$.  

The **minimum variance unbiased estimators (MVUE)** for $\mu$ and $\sigma^2$ are:
$$
\widehat{\mu}=\overline{X}=\frac{1}{n}\sum_{i=1}^n X_i \\
\widehat{\sigma^2}=S^2=\frac{1}{n-1}\sum_{i=1}^n (X_i - \overline{X})^2
$$

Use existing R functions to calculate the estimates of $\widehat{\mu}$ and $\widehat{\sigma^2}$.

```{r, collapse=T}
n <- length(x) #1000 
mu_hat <- sum(x) / n   #calculates mu_hat after taking the summation
mu_hat

sig <- (sum((x - mu_hat)^2))/(n-1) #calculates the summation of the differences first and then divides by length-1
sig
```

Do NOT use any existing R functions, write your own R code to calculate the estimates of $\widehat{\mu}$ and $\widehat{\sigma^2}$.

```{r, collapse=T}
sum <- 0
for(i in 1:length(x)){  #for loop to calculate the sum
sum <- sum + x[i]
}
mu_hat1 <- sum / length(x) #length(x) can be 1000 since its known
mu_hat1

sig1 <- 0
for(i in 1:length(x)){ 
sig1 <- sig1 + (x[i] - mu_hat1)^2    #summation of the differences 
}
sig1 <- sig1/ (length(x)-1)
sig1
```



```{r, collapse = T}
#par(mfrow=c(2,2))
z <- seq(0,8,length = 1000) #scale 
hist(quakes$mag, freq=F, xlab="Earthquake Magnitude", xlim = c(0,8)) # histogram - subfigure 1
lines(density(quakes$mag), col="red", lwd=3) # empirical density curve - subfigure 1
#prob.x <- dnorm(x, mu_hat, sig)
prob.z <- dnorm(z, mean = mu_hat, sd = sqrt(sig)) #normal distribution
lines(z, prob.z, lwd = 3, lty = 2, col = 'blue') # Plot the blue dashed line for normal distriubtion curve
```


**The Estimated normal curve is nearly similar to the empirical density curve. The magnitude obsverations follow a normal distribution and as you take a look at the Q-Q plot, one can conclude the same.** 



```{r}
# require(graphics) # load the graphics library

pairs(quakes, main = "Fiji Earthquakes, N = 1000", cex.main=1.2, pch=".")
```

Describe the output figure. 

**The figure for magnitude seems to form a series of vertical lines in the first image which contains a horizontal spacing in the second image, that spacing continues to widen into a fan-shaped pattern by the third image, and shows a roughly linear positive relationship by last image.** 


Do you think the `mag` variable and `stations` variables are positively correlated? Explain your answer.

**'Mag' and 'stations' are positively correlated because they both show similar pattern when they shift and change.** 



```{r, collapse = T}
cor_x <- cor(quakes$mag, quakes$stations) #correlation coefficient 
cor_x
cov_x <- cov(quakes$mag, quakes$stations) #covariance 
cov_x
plot(quakes$mag, quakes$stations, xlab= "Earthquake Magnitude", ylab = "Stations", pch = 16)
points(c(x[x < lwr], x[x > upr]), col=2, pch=23, lwd=5)
legend("topright",legend = c("Covariance: 0.85", "Correlation: 7.51", "x < lwr: 3.4 &  x > upr:5.8"), col=2, pch=23)

```



**FIJI is close by (east of) New Zealand and the Phillipines**  

```{r, collapse=T}
library(maps)
map()
points(quakes$long, quakes$lat, pch=".", col="red")
```


```{r, collapse=T}
long.mean <- mean(quakes$long)
lat.mean <- mean(quakes$lat)
#orient <- c(lat.mean, long.mean, 0)
xlim <- c(min(quakes$long)/2, max(quakes$long)*1.5)
ylim <- c(min(quakes$lat)-10, max(quakes$lat)+10)
map(database="world", xlim=xlim, ylim=ylim, col="grey80", fill=T)
```


```{r, collapse=T}
plot(quakes$long, quakes$lat, pch=".", col="red") # Plot the earthquake locations into the map 
outlierI <- which(x > upr) # Check outlier index
## [1]   15   17  152  558  753  870 1000
points(quakes$long[outlierI],quakes$lat[outlierI], pch=15, col="blue") #the outlier that we identified in part b
```


  
