---
title: 'Flights Visualization with Maps - Spring 2020'
author: 
- name: 'Sathwik Kesappragda'
  email: 
output:
  html_document:
    toc: yes
    toc_depth: 4
---


***

**Install necessary packages**

Note that you only need to install each package once. Then you can comment out the following installation lines.

``````{r}
#install.packages("maps")
#install.packages("mapproj")
```

***

**Load necessary packages**

```{r, collapse=T}
library(tidyverse) # for `ggplot2`, `dplyr`, and more

library(maps) # for map visualization
#library(mapdata)
library(mapproj) 

library(datasets) # for `state` data set
library(nycflights13) # for the 2013 NYC flights data set
```

***

```{r, collapse=T}
usa_map <- map_data("usa")
dim(usa_map)
glimpse(usa_map)
```

```{r}
ggplot(data = usa_map) + 
  geom_polygon(aes(x = long, y = lat, group = group)) + 
  coord_quickmap()
```


```{r, collapse = T}
#UCR plotted
ggplot(data = usa_map) + 
  geom_polygon(aes(x = long, y = lat, group = group), color = "blue", size = 1, fill = "red") + 
  geom_point(mapping = aes(x = -117.37 , y = 33.98 )) +
  coord_quickmap()

```

***

```{r, collapse=T}
states_map <- map_data("state")
dim(states_map)
glimpse(states_map)
```

```{r, collapse=T}
ggplot() + 
  geom_map(data = states_map, map = states_map,
           aes(map_id = region, fill = region), color="white") + 
  # geom_map() doesn't work in such a way that ggplot2 knows the extend of the map values, so you always have to expand the limits by hand yourself.
  expand_limits(x = states_map$long, y = states_map$lat) +
  coord_quickmap() +
  guides(fill = FALSE) # do this to leave off the color legend
```



```{r, collapse=T}
?state.x77
head(state.x77)

state_data <- as.data.frame(state.x77)
state_data$State <- tolower(rownames(state_data))
state_data %>% glimpse()
```

```{r, collapse=T}
ggplot(data = state_data, aes(map_id = State))+ 
  geom_map(map = states_map,
           aes(fill = Population)) + 
  expand_limits(x = states_map$long, y = states_map$lat) +
  coord_quickmap() 
```


```{r, collapse = T}
ggplot(data = state_data, aes(map_id = State))+ 
  geom_map(map = states_map,
           aes(fill = Income)) + 
  expand_limits(x = states_map$long, y = states_map$lat) +
  coord_quickmap() 
```

***


```{r, collapse=T}
head(state.center)
head(state.region)
```

```{r, collapse = T}

ggplot(data = state_data, aes(map_id = State))+ 
  geom_map(map = states_map, aes(fill = Population)) + #fill population was used to get a better outline of every state
  expand_limits(x = states_map$long, y = states_map$lat) +
  coord_quickmap() +
geom_point(mapping = aes(x=state.center$x, y=state.center$y,
                                      color = state.region,
                                      size = Population))  #just reproduced using the code from the previous exercise
``` 

***

```{r, collapse=T}
?flights # full documentation
glimpse(flights)
```


```{r, collapse = T}
not_canceled <- filter(flights, !is.na(dep_delay), !is.na(arr_delay))
dest_uniq <- 
  flights %>%
  select(dest) %>%
  distinct()

dest_uniq

num_flights <- 
  not_canceled %>%
  count(dest) 

num_flights

most_arrived <- not_canceled %>% 
  group_by(dest) %>%
  summarise(n = n()) %>%
  mutate(rank = rank(desc(n))) %>%
  filter(rank == 1)
most_arrived

least_arrived <- not_canceled %>% 
  group_by(dest) %>%
  summarise(n = n()) %>%
  mutate(rank = rank(n)) %>%
  filter(rank == 1)

least_arrived
```

*** 

```{r, eval=F}
#?airports
airports %>% glimpse()
per_dest<- 
  flights %>%
  select(dest) %>%
  distinct() 
  not_canceled %>%
  count(dest) 
left_join(per_dest, airports, by=c("dest"="faa"))
```


```{r, collapse = T}
per_dest<- 
  flights %>%
  select(dest) %>%
  distinct() 
  not_canceled %>%
  count(dest) 


flights_latlon<- not_canceled %>%
   inner_join(select(airports, origin = faa, origin_lat = lat, origin_lon = lon),
      by = "origin"
    ) %>%
   inner_join(select(airports, dest = faa, dest_lat = lat, dest_lon = lon),
     by = "dest"
   )
  left_join(per_dest, airports, by=c("dest"="faa"))


flights_latlon %>%
  ggplot(aes(
    x = origin_lon, xend = dest_lon,
    y = origin_lat, yend = dest_lat
  )) +
  borders("state") +
  geom_segment(arrow = arrow(length = unit(0.1, "cm"))) +
  coord_quickmap() +
  labs(y = "Latitude", x = "Longitude")


```



$$Y=2+3X+\epsilon$$
$$X \sim Uniform(A=-2, B=2)$$
$$\epsilon \sim N(\mu=0, \sigma=2)$$


```{r, collapse=T}
# Always remember to set the random seed for simulation studies
set.seed(167) # feel free to change 167 to your lucky number
X <- runif(100, -2, 2) ##simulating the data points
e <- rnorm(100, 0 , 2) ##simulate the error term N ~(0,2)
Y <- 2 + 3*X + e
#summary (y)
ggplot(mapping = aes(x = X, y = Y)) +
  geom_point() +
  geom_smooth(method = 'lm', se = F, col = "blue") 
```

***


the analytical solution of the least squares regression $Y = \beta_0 + \beta_1 X + \epsilon$ is

$$ 
\begin{aligned}
 \hat{\beta_1} & = 
 \frac{\sum_{i=1}^{n} (x_i - \bar{x})(y_i - \bar{y})}
 {\sum_{i=1}^{n} (x_i - \bar{x})^2} \\
 \hat{\beta_0} & = \bar{y} - \hat{\beta_1} \bar{x}
\end{aligned} 
$$

**After calculating the coefficients manually using the equations above, we see that the values match when compared to the values we get from the lm function.**  
```{r, collapse =T}
n <- length(X)
sxx <- sum(X^2) - sum(X)^2 / n
syy <- sum(Y^2) - sum(Y)^2 / n
sxy <- sum(X * Y) - (sum(X) * sum(Y)) / n

beta1 <- sxy / sxx
beta0 <- mean(Y) - beta1*mean(X)
beta1
beta0
# b1 <- cor(Y,X) * sd(Y)/sd(X)
# b0 <- mean(Y) - beta1*mean(X)
# b1
# b0

lin_reg <- lm(Y ~ X)
summary(lin_reg)
```

***



$$
\begin{aligned}
\text{RSE} = \hat{\sigma} &= \sqrt{\frac{1}{n-2}\text{RSS}} \\
&= \sqrt{\frac{1}{n-2} \sum_{i=1}^{n} ( y_i - \hat{y_i} )^2} 
\end{aligned}
$$ 

**The manually calculated residual standard error is not the same as the value we obtain from the lm function. The residual standard error shows the variability and how well the model fits the data.**  
```{r, collapse =T}
sse <- syy - sxy^2 / sxx
mse <- sse / (n - 2)
 
beta1.err <- sqrt(mse) / sqrt(sxx)
beta0.err <- sqrt(mse) / (sqrt(n) * sqrt(1 + (mean(X)^2 / (sum((X - mean(X))^2) / n))))
fitted <- X * beta1 + beta0
rse <- sqrt(sum((Y - fitted)^2) / (n - 2)) #rss in numerator, obtained from fitted 
rse 

lin_reg <- lm(Y ~ X)
summary(lin_reg)

```

***

$$
\begin{aligned}
\text{TSS} & = \sum_{i=1}^{n} (y_i - \bar{y})^2 \\
\text{RSS} & = \sum_{i=1}^{n} \left(y_i - \hat{f}(x_i) \right)^2 \\
R^2 & = \frac{\text{TSS}-\text{RSS}}{\text{TSS}} = 1 - \frac{\text{RSS}}{\text{TSS}} \\
\end{aligned}
$$
**The R-squared value is the same.**  
```{r, collapse = T}
r2 <- (syy - sse) /syy
r2
```

***


```{r, eval=F}
gg <- ggplot() +
  geom_abline(intercept = 2, slope = 3, col = "red", size = 1)
for ( i in 1:10) {
  ## add your simulation code here
  set.seed(i)
  X[[i]] <- runif(100,-2,2)
  e[[i]] <- rnorm(100,0,2)
  Y[[i]] <- 2 + 3*x[[i]] + e[[i]]
  local_data <- tibble(X = X[[i]], Y = Y[[i]])
  gg <- gg + 
    geom_smooth(data = local_data, mapping = aes(x=X, y= Y), col ="blue", size = 1) ## alternatively you can use geom_abline()
}

gg ## call gg at the end to plot it.
```


***